Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline

### 2014-04-07 paper.tex
Minor edit
Began adding responses to Reviewer 3's comments.

\begin{abstract}

We report how the impedance presented by a platinum electrode scales with the concentration of phosphate buffered saline (PBS).
We find that the constant phase element of the model scales with approximately the log of concentration, whereas the resistivity is inversely proportional.
Using a novel DC measurement technique we show that the Faradaic response of a platinum electrode, and thus the safe exposure limit, does not scale with concentration below 900\thinspace mV overpotential across a pair of electrodes.
We compare objective measurements made in saline to those made in the spinal cavity of live sheep. We comment upon the appropriateness of using PBS as a substitute for in-vivo measurements.
\end{abstract}

@@

\IEEEPARstart{T}{here} is considerable interest in the electrical modelling of electrodes. \cite{Cogan2008,Troy2006}
In \cite{Franks2005} a linearised model was presented, and in \cite{ScottSingle2013} a non-linear model
suitable for use with the SPICE circuit simulator was presented. The SPICE model usefully characterises an electrode in a given electrolyte with a small number of parameters.
One major reason for the interest in the electrical impedance of electrodes concerns the design of electronics intended for integration into pacemakers, cochlear implants, spinal cord stimulators, etc. The design of successful circuits depends upon a good understanding of external load impedance, while maximisation of battery life is linked to the use of electrodes whose impedance is well understood.
Having a compact model of an electrode in the appropriate electrolyte allows circuit designers to simulate their designs with valid loads. Knowing the effect upon the model of changes in the electrode or electrolyte would allow designers to anticipate circumstances that will affect the load seen by their circuits. Another appeal of a compact model stems from the fact that electrode characteristics are succinctly and objectively represented by the model parameters. These enable direct comparison of different electrodes, or the change in electrode properties over time. For example, in \cite{Kane13} changes in chronically-implanted electrodes are observed but there is no standard quantitative way of presenting the changes. Similarly, while there is a good understanding of what represents safe exposure, a compact model with parameters permits the prediction of the safety of any given stimulus regime.~\cite{Merrill05}

Electrodes in the laboratory are typically tested in a saline solution selected to mimic the circumstances in which they will operate when implanted.
In an implanted setting there is no reference electrode and many of the well established electrochemical measurement regimes do not apply. In this manuscript we measure the behaviour an electrode/electrolyte situation as it would be seen from an implant device.

@@

The electrode used in this work is a commercial linear array of eight platinum electrodes intended for SCS implantation called an ``Octrode''.~\cite{StJudeOctrode} A picture of an Octrode is presented in figure~\ref{fig:octrode}.
The electrical interface model of \cite{Franks2005} and \cite{ScottSingle2013} has two parts, a displacement branch and a Faradaic branch. While Scott and Single \cite{ScottSingle2013} incorporate a memristor into their model, it has been omitted here for simplicity. Its contribution is limited to situations where diffusion limited Faradaic reactions come into play, a situation that stimulator implants are designed to avoid.

@@

@@

\section{Resistive Network}
Six solutions ranging from 1.0X to 0.025X the concentration of a stock PBS solution have been examined.
The ingredients of that stock solution are given in table~\ref{tab:PBSrecipe}.
The pH level of the stock was measured, using a calibrated EDT Instruments BA-350 pH meter, to have a pH of 7.4. No pH adjustments were made to the derived solutions.

@@

\section{Displacement Parameters}

@@

\section{Faradaic Parameters}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Check that these are in the right order!!!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    {
        \color{Blue}
        We need this because it represents the resistance of the interface itself, separate from that of the fluid.
        We measure the impedance magnitude of the displacement currents at high frequency (flattened part of Fig. 5). For each concentration the impedance magnitude at 10kHz is higher than is predicted by the transimpedance measurements made in Fig. 3. The transimpedance measurments do not measure impedance across an interface that is passing current and therefore any voltage drop across an interface due to $R_{S}$ would vanish. The displacement measurements take the voltage across interfaces that are passing stimulus currents. These measurements see the potential drop across the interface due to the current flow via Ohms law.

@@

@@

    5. p 6, col 1, line 11. ``rate dependent on the log'' should be ``rate varies exponentially''. i.e. Tafel equation.
    {
        \color{blue}
        Correct, change made.
    }

@@

@@

>>>>

\begin{abstract}
We report how the impedance presented by a platinum electrode scales with the concentration of phosphate buffered saline (PBS).
{
    \color{blue}
    We measure this response in various solutions with an electrode array as is commonly used with spinal cord stimulation (SCS) implants. We match the parameters of a compact non-linear interface model to the measured response at each concentration.
}
We find that the constant phase element of the model scales with approximately the log of concentration, whereas the resistivity is inversely proportional.
Using a novel DC measurement technique we show that the Faradaic response of a platinum electrode, and thus the safe exposure limit, does not scale with concentration below 900\thinspace mV overpotential across a pair of electrodes.
We compare objective measurements made in saline to those made in the spinal cavity of live sheep. We comment upon the appropriateness of using PBS as a substitute for in-vivo measurements.
\end{abstract}

@@

\IEEEPARstart{T}{here} is considerable interest in the electrical modelling of electrodes. \cite{Cogan2008,Troy2006}
In \cite{Franks2005} a linearised model was presented, and in \cite{ScottSingle2013} a non-linear model
suitable for use with the SPICE circuit simulator was presented. The SPICE model usefully characterises an electrode in a given electrolyte with a small number of parameters.
One major reason for the interest in the electrical impedance of electrodes concerns the design of electronics intended for integration into pacemakers, cochlear implants, spinal cord stimulators, etc. The design of successful circuits depends upon a good understanding of external load impedance, while maximisation of battery life is linked to the use of electrodes whose impedance is well understood.

Having a compact model of an electrode in the appropriate electrolyte allows circuit designers to simulate their designs with valid loads. Knowing the effect upon the model of changes in the electrode or electrolyte would allow designers to anticipate circumstances that will affect the load seen by their circuits. Another appeal of a compact model stems from the fact that electrode characteristics are succinctly and objectively represented by the model parameters. These enable direct comparison of different electrodes, or the change in electrode properties over time. For example, in \cite{Kane13} changes in chronically-implanted electrodes are observed but there is no standard quantitative way of presenting the changes. Similarly, while there is a good understanding of what represents safe exposure, a compact model with parameters permits the prediction of the safety of any given stimulus regime.~\cite{Merrill05}

Electrodes in the laboratory are typically tested in a saline solution selected to mimic the circumstances in which they will operate when implanted.

In an implanted setting there is no reference electrode and many of the well established electrochemical measurement regimes do not apply. In this manuscript we measure the behaviour an electrode/electrolyte situation as it would be seen from an implant device.

@@

The electrode used in this work is a commercial linear array of eight platinum electrodes intended for SCS implantation called an ``Octrode''.~\cite{StJudeOctrode} A picture of an Octrode is presented in figure~\ref{fig:octrode}.

The electrical interface model of \cite{Franks2005} and \cite{ScottSingle2013} has two parts, a displacement branch and a Faradaic branch. While Scott and Single \cite{ScottSingle2013} incorporate a memristor into their model, it has been omitted here for simplicity. Its contribution is limited to situations where diffusion limited Faradaic reactions come into play, a situation that stimulator implants are designed to avoid. {\color{blue} Modelling Faradaic current onset allows engineers to predict and avoid such situations in-vivo. For this reason the measurements and model fitting are mostly concerned with capturing only the onset of Faradaic conduction. The model is not intended to model the behaviour of an electrode array operating in a Faradaic conduction mode.}

@@

{
    \color{blue}
    The remainder of this manuscript is broken into five sections. Section \ref{sect:resistorMesh} details the measurement of the inter-electrode resistances and the generation of a resistor network to replicate these resistances.
    Sections \ref{sect:cpe} and \ref{sect:faradaic} are concerned with the measurement and parameter fitting of displacement currents and Faradaic currents, respectively, across the electrolyte interface.
}

@@

\section{{\color{blue}Bulk Resistivity / Resistor} Network }
\label{sect:resistorMesh}
Six solutions ranging from 1.0X to 0.025X the concentration of a stock PBS solution have been examined.
The ingredients of that stock solution are given in table~\ref{tab:PBSrecipe}.
The pH level of the stock was measured, using a calibrated EDT Instruments BA-350 pH meter, to have a pH of 7.4. No pH adjustments were made to the derived solutions.

@@

\section{Displacement Parameters {\color{blue}/ CPE}}
\label{sect:cpe}

@@

\section{Faradaic Parameters {\color{blue}/ Diodes}}
\label{sect:faradaic}

@@

@@

    {
        \color{blue}
        We need this because it represents the resistance of the interface itself, separate from that of the fluid.
        We measure the impedance magnitude of the displacement currents at high frequency (flattened part of Fig. 5). For each concentration the impedance magnitude at 10kHz is higher than is predicted by the transimpedance measurements made in Fig. 3. The transimpedance measurements do not measure impedance across an interface that is passing current and therefore any voltage drop across an interface due to $R_{S}$ would vanish. The displacement measurements take the voltage across interfaces that are passing stimulus currents. These measurements see the potential drop across the interface due to the current flow via Ohms law.

@@

    {
        \color{blue}

@@

    {
        \color{blue}
        The RC pairs are generated via a Python script for a given frequency range. Simulations in this manuscript used RC pairs spanning 1.0 nHz to 1.0 GHz spaced 3 per decade. Adding more RC pairs than necessary came at negligible computation cost and having the circuited constructed in an automated fashion to satisfy the given frequency range adds no extra work.
    }

    5. p 6, col 1, line 11. ``rate dependent on the log'' should be ``rate varies exponentially''. i.e. Tafel equation.

    {
        \color{blue}
        Correct, change made.
    }

@@

    {
        \color{blue}
        Of course, thank you.
    }

@@

    {
        \color{blue}
        Heading expanded slightly to include model elements.
    }
